An upper bound to gas delivery via pressure-swing adsorption in nanoporous materials

### 2020-05-16 gas-adsorption/gas-ads-appendix.tex
fixed the appendix; now to add the figure.

\section{The real-substrate analog of $\V$ is $\gst$}
\label{sec:phi-is-delta-g}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{g_st_piston_2.png}
    \caption{Cartoon illustrating the definition of $\gst$ and a
      hypothetical experiment to measure it with a piston and
      removable barrier that separates two equal volumes, one with a
      porous material (bottom) and one with free space (top). The
      initial state (left) is a bulk gas phase at density $\rho$ and
      temperature $T$. An infinitely-thin, removable barrier separates
      the gas from the porous material (bottom). The barrier is then
      slowly removed, and the piston slowly pushes all gas molecules
      into the porous material in an isothermal process. The final
      state (right) is an adsorbed gas phase at density $\rho$ and
      temperature $T$. The quantity $\gst$ follows from \davidsays{the heat released?}.}
    \label{fig:delta-G-cartoon}
\end{figure}

The parameter describing our idealized substrate is $\V$, the spatially uniform potential felt by a gas molecule adsorbed in the idealized substrate. 
A natural question is how this potential relates to the properties of real substrates. 
The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the gas (see Sec.~\ref{sec:V_shifts_chem_pot}).
Because our ideal substrate shifts the chemical potential of the gas molecules by providing a spatially uniform potential energy field, the entropy of the gas in the ideal substrate is equal to the entropy of the gas in its bulk state at the same density and temperature.
In contrast, a real substrate provides a non-spatially uniform potential. Consequently, the entropy of the gas inside a real substrate is \emph{not} equal to the entropy of the bulk gas at the same density and temperature. 
Therefore, the parameter analogous to $\V$ in a real substrate will involve both energy and entropy.
The real-substrate analog of $\V$ is the shift of molar Gibbs free energy provided by the substrate, specifically an \emph{isosteric} (or constant-density) shift of the Gibbs free energy:
\begin{equation}
   \gst(\rho, T) \equiv
   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
  \label{eq:g_st}
\end{equation}
The isosteric Gibbs free energy difference $\gst$ is the difference in molar Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$ does \emph{not} correspond to a change in the molar Gibbs free energy as a molecule is adsorbed, which is zero under conditions of coexistence. 
The quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal substrate because it is the chemical potential shift needed to impose on the bulk gas in coexistence with the real substrate to achieve the same density as in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to  measure $\gst$ via a piston with a removable partition that separates a volume of free space from the same volume of substrate. Note that $\gst$ is a property of both the substrate and the identity of the gas. Because real substrates offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho, T)=\V$. Consequently, throughout this article, we show $\gst(\rho, T)$ for real substrates at both conditions relevant to gas storage and delivery, $\pfull$ and $\pempty$. 
% In the initial state, the partition isolates the bulk gas exhibiting density $\rho$ and temperature $T$. Then, the partition is removed and the piston pushes down on the gas so that the gas slowly adsorbs into the MOF in an isothermal process. \red{isothermal rite?} In the final state, the piston has pushed all gas into the MOF, resulting in an adsorbed gas phase with the same density $\rho$ and temperature $T$ as the initial state. The isosteric Gibbs free energy of adsorption, $\gst$, then follows from \red{is it the amount of heat released?}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
    \caption{Relationship between $\gst$ and the isosteric heat
      $q_{st}$ for several prominent adsorbents at room
      temperature. The dots represent the properties of methane
      adsorption at 5~bar and 65~bar. The $+$'s represent the
      properties of hydrogen adsorption at 5~bar and
      100~bar. \davidsays{Add citations.}}
    \label{fig:qst-vs-delta-G}
\end{figure}

@@

\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding to a full and empty density. The deliverable capacity of the MOF is equal to the difference between the full and empty density.
Examining Fig.~\ref{fig:methane-gst}, we see that a significant variety of known experimental $\Delta g_\text{st}$ curves are monotonic.  Furthermore, our qualitative argument in Section~\ref{sec:upper-bound} suggests that this function \emph{should} be monotonically increasing for rigid substrates, as increasing the density of gas forces some of the gas to reside in higher-energy sites.  If this is always the case, the deliverable capacity will always be bounded by our theoretical upper bound, even for a value of $\V$ that is not optimal.

@@

We now consider what happens if we introduce a diffusive connection between two boxes that are initially at the same density, for instance by connecting them with a tube.  If the chemical potential of the two boxes are equal, their density will be unchanged, otherwise gas will flow into the box with lower chemical potential.  The difference in chemical potential is equal to the difference
\begin{align}
   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
   \\
   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
\end{align}
Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at their initial density after they are connected.  Thus the deliverable capacity of the MOF will be greater than the deliverable capacity of the model with uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$ does \emph{not} monotonically increase.  By the same token, if we consider the case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable capacity than our model, the MOF must have less residual gas, which means that gas must spontaneously flow from the low-density MOF to the box with potential $\V$, which means that $\gst(\rhoempty)>\V$.  So once again exceeding our upper bound requires a material with a $\gst(\rho)$ that does not increase monotonically.

@@

\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
\begin{figure}
    \centering
    FIGURE HERE A B C D
    \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$. Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
    \label{fig:hydrogen-77-D}
\end{figure}

>>>>

\section{The real-substrate analog of $\V$ is $\gst$}
\label{sec:phi-is-delta-g}
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{g_st_piston_2.png}
    \caption{Cartoon illustrating the definition of $\gst$ and a
      hypothetical experiment to measure it with a piston and
      removable barrier that separates two equal volumes, one with a
      porous material (bottom) and one with free space (top). The
      initial state (left) is a bulk gas phase at density $\rho$ and
      temperature $T$. An infinitely-thin, removable barrier separates
      the gas from the porous material (bottom). The barrier is then
      slowly removed, and the piston slowly pushes all gas molecules
      into the porous material in an isothermal process. The final
      state (right) is an adsorbed gas phase at density $\rho$ and
      temperature $T$.}
    \label{fig:delta-G-cartoon}
\end{figure}

The parameter describing our idealized substrate is $\V$, the spatially uniform
potential felt by a gas molecule adsorbed in the idealized substrate. A natural
question is how this potential relates to the properties of real substrates.
The effect of $\V$ in our model is to shift the chemical potential $\mu$ of the
gas (see Sec.~\ref{sec:V_shifts_chem_pot}). Because our ideal substrate shifts
the chemical potential of the gas molecules by providing a spatially uniform
potential energy field, the entropy of the gas in the ideal substrate is equal
to the entropy of the gas in its bulk state at the same density and
temperature. In contrast, a real substrate provides a non-spatially uniform
potential. Consequently, the entropy of the gas inside a real substrate is
\emph{not} equal to the entropy of the bulk gas at the same density and
temperature. Therefore, the parameter analogous to $\V$ in a real substrate
will involve both energy and entropy. The real-substrate analog of $\V$ is the
shift of molar Gibbs free energy provided by the substrate, specifically an
\emph{isosteric} (or constant-density) shift of the Gibbs free energy:
\begin{equation}
   \gst(\rho, T) \equiv
   %\frac{G_{\text{ads}}(\rho, T) - G_{\text{gas}}(\rho, T)}{N}\\ 
    \mu_{\text{ads}}(\rho, T) - \mu_{\text{gas}}(\rho, T).
  % \gst(\rho, T) &\equiv g_{\text{gas}}(\rho, T) - g_{\text{ads}}(\rho, T)
  \label{eq:g_st}
\end{equation}
The isosteric Gibbs free energy difference $\gst$ is the difference in molar
Gibbs free energy ($=$ chemical potential) between the adsorbed gas system and
the bulk gas \emph{with the same density of gas molecules}. The quantity $\gst$
does \emph{not} correspond to a change in the molar Gibbs free energy as a
molecule is adsorbed, which is zero under conditions of coexistence. The
quantity $\gst$ in a real substrate is a direct analog to $\V$ in our ideal
substrate because it is the chemical potential shift needed to impose on the
bulk gas in coexistence with the real substrate to achieve the same density as
in the substrate (compare with eqn.~\ref{eq:xi_vs_xi0}).
Figure~\ref{fig:delta-G-cartoon} illustrates a hypothetical experiment to
measure $\gst$ via a piston with a removable partition that separates a volume
of free space from the same volume of substrate. Note that $\gst$ is a property
of both the substrate and the identity of the gas. Because real substrates
offer a \emph{non}-spatially uniform potential, $\gst(\rho, T)$ is a function
of $\rho$ and $T$, unlike our ideal, homogenous substrate where $\gst(\rho,
T)=\V$. Consequently, throughout this article, we show $\gst(\rho, T)$ for real
substrates at both conditions relevant to gas storage and delivery, $\pfull$
and $\pempty$.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{qst-vs-delta-G}
    \caption{Relationship between $\gst$ and the isosteric heat
      $q_{st}$ for several prominent adsorbents at room
      temperature (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}). The dots represent the properties of methane
      adsorption at 5.8~bar and 65~bar. The $+$'s represent the
      properties of hydrogen adsorption at 5~bar and
      100~bar.}
    \label{fig:qst-vs-delta-G}
\end{figure}

@@

\section{An upper bound when $\gst(\rho)$ is monotonic}\label{sec:monotonic}
Every MOF has a $\Delta g_\text{st}$ at a full and empty pressure corresponding
to a full and empty density. The deliverable capacity of the MOF is equal to
the difference between the full and empty density. Examining
Fig.~\ref{fig:methane-gst}, we see that a significant variety of known
experimental $\Delta g_\text{st}$ curves are monotonic. Furthermore, our
qualitative argument in Section~\ref{sec:upper-bound} suggests that this
function \emph{should} monotonically increase for rigid substrates, as
increasing the density of gas causes some of the gas to reside in higher-energy
sites. If this is always the case, the deliverable capacity will always be
bounded by our theoretical upper bound, even for a value of $\V$ that is not
optimal.

@@

We now consider what happens if we introduce a diffusive connection between two
boxes that are initially at the same density, for instance by connecting them
with a tube. If the chemical potential of the two boxes are equal, their
density will be unchanged, otherwise gas will flow into the box with lower
chemical potential. The difference in chemical potential is equal to the
difference
\begin{align}
   \gst(\rho) - \V &= \left(\mu_{\text{MOF}}(\rho) - \mu_{\text{gas}}(\rho)\right)
   - \left(\mu_{\V}(\rho) - \mu_{\text{gas}}(\rho)\right)
   \\
   &= \mu_{\text{MOF}}(\rho) - \mu_{\V}(\rho).
\end{align}
Thus if $\gst(\rhoempty) = \V$ then the ``empty'' containers will remain at
their initial density after they are connected. Thus, the deliverable capacity
of the MOF will be greater than the deliverable capacity of the model with
uniform potential $\V$ if and only if $\gst(\rhofull)<\V$, i.e. if $\gst(\rho)$
does \emph{not} monotonically increase. By the same token, if we consider the
case where $\gst(\rhofull) = \V$, then to achieve a greater deliverable
capacity than our model, the MOF must have less residual gas, which means that
gas must spontaneously flow from the low-density MOF to the box with potential
$\V$, which means that $\gst(\rhoempty)>\V$. Once again, exceeding our upper
bound requires a material with a $\gst(\rho)$ that does not increase
monotonically.

@@

\section{Cryogenic hydrogen storage}\label{sec:cryo-hydrogen}
\begin{figure}
    \centering
    FIGURE HERE A B C D
    \includegraphics[width=0.95\columnwidth]{hydrogen-77-n-vs-G}
    \caption{Deliverable capacity of hydrogen at liquid nitrogren temperature as a function of the attractive Gibbs free energy $\gst$. Experimental deliverable capacities for several MOFs (data from Ref.~\cite{garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
    \label{fig:hydrogen-77-D}
\end{figure}

### 2020-05-16 gas-adsorption/gas-ads-body.tex
fixed the appendix; now to add the figure.

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{hydrogen-298-n-vs-G}
    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.  Experimental deliverable capacities for several MOFs are shown along with the experimental values for $\gst$ at the empty and full pressures shown as x's connected by a line.}
    \label{fig:hydrogen-298-D}
\end{figure}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.95\columnwidth]{hydrogen-298-n-vs-G}
    \caption{Deliverable capacity of hydrogen at room temperature as a function of the attractive Gibbs free energy $|\gst|$.  Experimental deliverable capacities for several MOFs (data from Ref.~\cite{mason2014evaluating, garcia2018benchmark}) are shown along with the experimental values for $\gst$ at the empty and full pressures shown as $+$'s connected by a line.}
    \label{fig:hydrogen-298-D}
\end{figure}
